---
title: 'R for data science '
author: "Thulani"
output: html_notebook
---

## Insert: select to code in R
loading libraries
```{r}
library(ggplot2)

```

```{r}
mpg
```

Setting working directory
```{r}
setwd("C:/Users/thula/Desktop/Workshop_datascience")
getwd()
list.files()
```

### variables
```{r}
num1<-45
num1

num2<-60
num2
num1+num2

```

### Functions
```{r}
sum_function = function(x=NULL,y=NULL){
  sum(x,y)
}
sum_function(2,5)
```
#sum_function = function(...) #when we don't know what variablws to be used in function

### Vectors
```{r}
vec=c(1:10)
vec
#repeat 1 for 10 times
rep(1,10)
```
```{r}
seq(1,100,2)
```

### Dataframes
```{r}
df=data.frame(index=c(1:10),name=c(letters[1:10]))
df
head(df,n=4)
dim(df)
str(df)
class(df)
levels(df$name)
#all rows of column 1
df[,1]
#mean of index column 
mean(df$index)
mean(df[,1])

```

### Import a file: tab seperated bed file
```{r}
df_tf = read.table(file="TF_GRCh37.bed", sep="\t")
df_tf

#to load large data
library(data.table)
df_tf=fread("TF_GRCh37.bed")
df_tf
colnames(df_tf)=c("chr","start","end","TF","score")
colnames(df_tf[,5])="score2"
```

### Using ggplot 
```{r}
#Processed rectangular data set: transformed
mpg
class(mpg)
ggplot(data=mpg)+
  geom_point(aes(x=displ,y=hwy))+
  theme_bw()
```

### Apply colours
```{r}
ggplot(data=mpg)+
  geom_point(aes(x=displ,y=hwy,color=class))+
  theme_bw()
```

### Size and shape
```{r}
#size
ggplot(data=mpg)+
  geom_point(aes(x=displ,y=hwy,color=class, size=trans))+
  theme_bw()

#shape
ggplot(data=mpg)+
  geom_point(aes(x=displ,y=hwy,color=class, shape=manufacturer))+
  theme_bw()
```


### To find patterns in data
```{r}
ggplot(data=mpg)+
  geom_smooth(aes(x=displ,y=hwy,group=drv,color=drv))+
  theme_bw()
```

### Geom_point and smooth together
```{r}
ggplot(data=mpg)+
  geom_point(aes(x=displ,y=hwy,color=class))+
  geom_smooth(aes(x=displ,y=hwy,group=drv))+
  theme_bw()
```

### Categorizing cars depending on fuel types
```{r}
ggplot(data=mpg)+
  geom_bar(mapping=aes(x=mpg$fl),data=mpg)+
  theme_bw()
```

### City mileage for each data type
```{r}
ggplot(data=mpg)+
  geom_bar(aes(x=cty,color=fl,fill=fl))
  theme_bw()
```

### Importing hg19 data
```{r}
library(data.table)
df_tf2=fread("2009GRCh37_hg19.bed")
TF=factor(df_tf2$V4)
g=ggplot(data=df_tf2)+
  geom_bar(aes(x=df_tf2$V4),position = "identity")
  theme_bw()
g

X=as.data.frame(table(df_tf2$V4))
X
X=X[order(-X$Freq),]
X
X[X$Freq == max(X$Freq),]
```

### Analysis using dplyr package
### Grammar of data manipulation:  improves readability
```{r}
library(dplyr)
dt_tf2=as.data.frame(table(df_tf2$V4))

#pipes - %>%
df_tf3=df_tf2$V4 %>% table %>% as.data.frame %>% head
df_tf3
df_tf2 %>% group_by(V4) %>%  summarise(freq = n())
```

### Subsetting 
```{r}
Chr1=df_tf2 %>% subset(V1=="chr1" ) 
Chr2=df_tf2 %>% subset(V1=="chr2") 
Chr2
#column names should be equal
rbind(Chr1,Chr2)
#number of columns should be equal
cbind(Chr1,Chr2)
```

### Add length of binding site of TF as a column
```{r}
length=data.frame((Chr1$V3)-(Chr1$V2))
length
newdf=cbind(Chr1$V1,length)
newdf
len=length$index=c(1:dim(length)[1])
len
```

### To merge two files with same unique id column
```{r}
#like joins in db
#x.all= T -> all rows of X will be merges(outer join)
merge(x=Chr1, y= length, by.x = Chr1$,by.y=length$ ,)
```

### mutate

